---
title: 'in situ vignette: data checking and calibration'
author: "Christina Bonsell"
date: "April 9, 2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{in situ vignette: data checking and calibration}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Some intro text

First let's load in some handy packages
```{r}
library(insitu)
library(RCurl) #For loading in data directly from online repositories
library(tidyverse) #we're going to work in the tidyverse for data manipulation and plotting
library(cowplot)
```

#Calculate salinity from conductivty and temperature ising the UNESCO algorithm

```{r}
calculate_salinity(25.50,1)
```



#easily read in files from either RBR or starOddi CTDs, or Lowell Tilt Current Meters  `r ?import_data`

***

#How to Perform quality control on CTD datasets, with particular focus on Arctic nearshore systems

###Look at the data

First, read in some data (this is Beaufort Lagoon Ecosystem LTER data from 2018-2019)

```{r}
urlCSV <- getURL("https://portal-s.edirepository.org/nis/dataviewer?packageid=knb-lter-ble.3.7&entityid=534373640a661d8ed0bfedc52479a133", timeout = 200)
txtCSV <- textConnection(urlCSV)
tempsal_data <- read.csv(txtCSV, stringsAsFactors = F) %>% 
  filter(station=="KALD1") %>% 
  mutate(date_time=as.POSIXct(date_time, format="%m/%d/%Y %H:%M"))
close(txtCSV)

head(tempsal_data)
str(tempsal_data)
```


Simple timeseries plots
```{r}
ggplot(tempsal_data, aes(date_time, temperature))+
  geom_line()

ggplot(tempsal_data, aes(date_time, conductivity))+
  geom_line()
```

Some of this conductivity data is clearly bad (look at that June-July weirdness)! We need to calibrate the data and remove "bad" points.


#Calibrate the data based on readings from a more recently-calibrated sensor

read in data from YSI sonde measurements after deployment and just before retrieval of the CTD
```{r}
urlCSV <- getURL("https://portal-s.edirepository.org/nis/dataviewer?packageid=knb-lter-ble.3.7&entityid=6d327300d4893e4f932220fecf49c507", timeout = 200)
txtCSV <- textConnection(urlCSV)
ysi<- read.csv(txtCSV, stringsAsFactors = F) %>% 
  filter(station=="KALD1") %>% 
  mutate(date_time=as.POSIXct(substr(date_time,1, 16), format="%Y-%m-%dT%H:%M"))#datetime is in ISO 8601, but inports as character
close(txtCSV)

head(ysi)

```

Calibrate the data based on linear relationship between these two "anchor" points. Note that this will also work with >2 anchor points, for which the adjustment will be based on the linear relationship between each sequential pair of anchor points. If data exists before or after anchor points, the linear relationship will use the first (in the before case), or last (in the after case) reading from the raw data. 

```{r}
tempsal_data_calibrated <- calibrate_data(tempsal_data, ysi, raw= "temperature", cal_by = "temperature_C")

tempsal_data_calibrated  <- calibrate_data(tempsal_data_calibrated, ysi, raw= "conductivity", cal_by = "conductivity_uS_cm")

head(tempsal_data_calibrated)

ggplot(tempsal_data_calibrated, aes(date_time, temperature))+
  geom_line()+
  geom_line(aes(date_time, temperature_calibrated), color="blue")

ggplot(tempsal_data_calibrated, aes(date_time, conductivity))+
  geom_line()+
  geom_line(aes(date_time, conductivity_calibrated), color="blue")
```

Ok, now that the data is calibrated, let's calculate and plot salinity. We're going to to this the "tidyverse" way
```{r}
tempsal_data_calibrated <- tempsal_data_calibrated %>%
  mutate(salinity_calibrated=calculate_salinity(conductivity_calibrated, temperature_calibrated))

ggplot(tempsal_data_calibrated, aes(date_time, salinity_calibrated))+
  geom_line()
```

Clearly some of that data is still bad, so let's flag it...


#Identify "bad" data based on the freezing line from temperature/salinity plots
Temperature/salinity plots can easily tell you when something is amiss in your data. This often occurs because the conductivity reading if off due to obstruction by sediments or fouling (for instruments with a conductivity cup), or by nearby ice or other conductive materials (for RBR "donuts"). In cold regions, data below the freezing line is expecially suspect.

```{r}
plot_tempsal(tempsal_data_calibrated, "temperature_calibrated", "salinity_calibrated", "date_time", plottitle="example")
```

The straight lines for points in the temp vs sal plot indicate intstrument error. We can remove the points below the freezing line programatically.

```{r}
flag_salinity
```


#You can also manually identify "bad" data that doesn't get picked up from the freezing line algorithm
